A 3d curvilinear skeletonization algorithm with application to path tracing

### 2012-09-13 application_to_path_tracing.tex
Nouvelles corrections
Correction + modif partie path tracing + images

\section{Application to path tracing}

Path tracing is a global illumination algorithm that is able to render photo-realistic images of a virtual scene viewed by a camera. The idea behind the algorithm is very elegant: for each pixel of the image, throw rays through the pixel that will bounce on the scene to accumulate lighting exchanges. In order to get fast convergence the rays must be efficiently distributed through the scene, toward the lights. Our idea is to use a curvilinear skeleton of the void of the scene to guide the rays in the correct direction. We start by defining some useful concepts related to path tracing and we present our skeleton based path tracing algorithm.

@@

Let $O$ be the origin of the camera in the scene. For each pixel $P$ of the final image we search for the nearest intersection point $x$ of the ray $(O, \overrightarrow{OP} = -\Theta)$ with the scene. At this point, to obtain the luminosity, we must solve the rendering equation \cite{Ka86} which is a recursive integral equation. The integrand of the equation contains the radiance function that we must compute:

\begin{align*}
L(x \rightarrow \Theta) = L_e(x \rightarrow \Theta) + \int_{\Phi \in \Omega_x} f_r(x, \Theta \leftrightarrow \Phi) L(r(x, \Phi) \rightarrow -\Phi) cos(N_x, \Phi) d\omega_\Phi
\end{align*}

$L$ is the radiance (a radiometric quantity that represents luminosity) from point $x$ toward direction $\Theta$. $L_e$ is the emitted radiance that is not zero only on light sources. $r(x, \Phi)$ is the nearest visible point from $x$ in the direction $\Phi$ of the hemisphere $\Omega_x$. $N_x$ is the surface normal at point $x$. The $f_r$ function is called the \myem{bidirectional reflectance function} (BRDF) and expresses the exchange of radiance between the incoming direction $\Phi$ and the output direction $\Theta$. Intuitively it represents the material properties of the surface at point $x$. A mirror does not reflect light in the same manner as a wall does and the BRDF describes that kind of behavior.

This equation expresses an intuitive idea: the radiance $L$ produced by the point $x$ in the direction $\Theta$ is the sum of the emitted radiance and the reflected radiance.  The reflected radiance is computed by integrating the incoming radiance scaled by a factor $f_r(x, \Theta \leftrightarrow \Phi) cos(N_x, \Phi)$ over all the possible directions of the hemisphere around $x$.

The most commonly used method to compute the integral is the Monte-Carlo integration that provides the following estimator:
\begin{equation*}
\langle L_r(x \rightarrow \Theta) \rangle = \frac{f_r(x, \Theta \leftrightarrow \Phi) L(r(x, \Phi) \rightarrow -\Phi) cos(N_x, \Phi)}{p(\Phi)}
\end{equation*}

where $p$ is a probability density function (pdf) that is used to sample $\Phi$. This estimator is unbiased, meaning that the expected value $E[\langle L_r(x \rightarrow \Theta) \rangle]$ is equal to $L_r(x \rightarrow \Theta)$. The variance of the estimator expresses its quality and depends on the choosen pdf $p$. Actually the best is to choose a pdf that matches the shape of the function to integrate (ie. give high density to samples that have high values for the function and low density to samples that have low values).

This strategy of choosing an adapted pdf is called \myem{importance sampling} and is used in global illumination to improve the convergence speed of the algorithms.

The algorithm to compute $L$ is as follow: first we sample a random direction $\Phi$ based on the pdf $p$. Then we apply the estimator that calls $L$ recursively to compute $L_r$. Finally we return the sum of the emitted radiance $L_e$ and the result for $L_r$. The recursion stops when a maximal number of bounces has been reached.

This computation is done multiple times for each pixel (at different positions over the pixel). We average the results and get an estimation of the mean radiance passing through the pixel in direction of the camera's origin.

In the radiance computation the $\Phi$ direction is sampled on $\Omega_x$ based on the pdf $p$. A bad pdf picks directions that don't reach the light before the end of the recursion and results in a lot of noise in the final image. We will discuss the choice of $p$ in the next section.

@@

The key idea behind our method is that light travels in the void of the scene. As stated in the introduction, a curvilinear skeleton of the void gives us enough information to compute some importance directions where light comes from. Given these directions we can construct a efficient pdf $p_{skel}$ and guide our rays by sampling the hemispheres with $p_{skel}$.

@@

@@

The skeleton of the void is computed with our algorithm (section \ref{sec::skeletonization}) and is converted to a graph (the nodes are 3D positions and the edges represent the topology of the scene), given as an input of our path tracing. We then compute a set of points called \myem{importance points}. These points will be used to sample $\Omega_x$ in the path tracing algorithm. For each node of the skeleton $n$, one importance point $imp_n$ is computed.

Let $L$ be a light source of the scene ($L \in S_L$) and $n_L$ the nearest node that is visible by $L$. We first compute a tree of shortest paths along the skeleton with $n_L$ as root. We use the Dijkstra algorithm with a distance based on visibility from $n_L$: We put $d(n, m) = 1$ if $m$ is visible from $n_L$ and $d(n, m) = 10$ if not. It results that the paths that are enlighted by $L$ will have a shorter distance and will guide us to the light faster than the dark ones.

@@

We observe a decrease of the noise for our algorithm, especially in dark areas. The MSE values are lower for the top pictures, meaning that we converge faster to the reference image than standard path tracing does.

>>>>

\section{Application to path tracing}
\label{sec::appli_pathtrace}

In this section, we explain how to use the curvilinear skeleton of the voids of the scene in order to enhance the path tracing algorithm. We start by defining some basic elements related to path tracing and we present our skeleton based path tracing algorithm.

@@

Let $O$ be the origin of the camera in the scene. For each pixel $P$ of the final image, let $x$ be the nearest intersection point of the ray $(O, \overrightarrow{OP} = -\Theta)$ with an element of the scene. To obtain the luminosity, we must solve the rendering equation \cite{Ka86} which is a recursive integral equation. The integrand of the equation contains the radiance function that we must compute:

\begin{align*}
L(x \rightarrow \Theta) = L_e(x \rightarrow \Theta) + \int_{\Phi \in \Omega_x} f_s(x, \Theta \leftrightarrow \Phi) L(r(x, \Phi) \rightarrow -\Phi) d\omega_\Phi
\end{align*}

$L$ is the radiance (a radiometric quantity that represents luminosity) from point $x$ toward direction $\Theta$. $L_e$ is the emitted radiance (non null only on light sources). The point $r(x, \Phi)$ is the nearest visible point from $x$ in the direction $\Phi$ of the hemisphere $\Omega_x$. The $f_s$ function expresses how much luminosity is exchanged, on point $x$, between an incoming ray $\Phi$ and the outgoing ray $\Theta$. Intuitively it represents the material properties of the surface at point $x$.

This integrand expresses an intuitive idea: the reflected radiance from $x$ towards the camera is the result of computing all incoming luminosity on $x$ scaled by a factor which depends on the material on $x$ and the angle of the ray going from the camera to $x$. The most common method used to estimate the integral is the Monte-Carlo integration which provides the following:

\begin{equation*}
\langle L_r(x \rightarrow \Theta) \rangle = \frac{f_s(x, \Theta \leftrightarrow \Phi) L(r(x, \Phi) \rightarrow -\Phi)}{p(\Phi)}
\end{equation*}

where $p$ is a probability density function (pdf) that is used to sample $\Phi$. This estimator is unbiased, meaning that the expected value $E[\langle L_r(x \rightarrow \Theta) \rangle]$ is equal to $L_r(x \rightarrow \Theta)$. The variance of the estimator expresses its quality and depends on the choosen pdf $p$. The best choice is a pdf that matches the shape of the function to integrate (ie. give high density to samples that have high values for the function and low density to samples that have low values).
The strategy of choosing an adapted pdf is called \myem{importance sampling} and is used in global illumination to improve the convergence speed of the algorithms.

The algorithm computes $L$ as follow: first, it chooses one random direction $\Phi$ based on the pdf $p$. Then, it applies the estimator by calling $L$ recursively to compute $L_r$ (shoots a new ray in the direction $\Phi$, computes the new hit point with the scene, and compute the new radiance at this point). Finally, it returns the sum of the emitted radiance $L_e$ and the result for $L_r$. The recursion stops either when a light or a maximal number of bounces has been reached.
%ASK: euh, c'est L ou L_r%
%c'est bon de dire que ça s'arrete au light%

This computation is done multiple times for each pixel (at different positions over the pixel). We average the results and get an estimation of the mean radiance passing through the pixel and heading towards the camera.
%at different positions over the pixel  ???%
A bad pdf would lead in picking directions that don't reach the light before the end of the recursion, and produce results with a lot of noise in the final image. In the next part, we explain how to produce a pdf based on the curvilinear skeleton.

@@

As stated in the introduction, a curvilinear skeleton of the void (with some preprocessing achieved on it) gives us information on which directions the light comes from. Given these directions, we can build a efficient pdf $p_{skel}$ and guide our rays by sampling the hemispheres with $p_{skel}$.

@@

%a revoir, je ne comprends pas tout

@@

The skeleton of the voids of the scene is computed using Alg.~\ref{algo::3d_curvskel} and is converted to a graph (the nodes are the 0-faces and the edges are the 1-faces). We then compute a set of \myem{importance points}: these points will be used to sample $\Omega_x$ in the path tracing algorithm. For each node of the skeleton $n$, one importance point $imp_n$ is computed. Intuitively, the importance point associated to $n$ gives us a direction to follow in order to find a light source.
%TODO a verifier quand meme

%La, pas clair !!!
Let $L$ be a light source of the scene ($L \in S_L$) and $n_L$ the nearest node that is visible by $L$. We first compute a tree of shortest paths along the skeleton with $n_L$ as root. We use the Dijkstra algorithm with a distance based on visibility from $n_L$: We put $d(n, m) = 1$ if $m$ is visible from $n_L$ and $d(n, m) = 10$ if not. It results that the paths that are enlighted by $L$ will have a shorter distance and will guide us to the light faster than the dark ones.

@@

We observe that images produced by our algorithm have less noise than the ones produced by the regular path tracing (for a same number of rays shot per pixel). The MSE values are lower for the top pictures, meaning that we converge faster to the reference image than standard path tracing does.

### 2012-09-13 skeleton_and_pathtracing.tex
Correction + modif partie path tracing + images

\def\figpath#1{../../Images/EPS/#1}

@@

\begin{abstract}
What we propose is really great...
\end{abstract}

@@

On a single execution of the main loop of Alg.~\ref{algo::pardircollapse_v0}, only faces located on the border of the complex are removed (l.~\ref{pc:07}). Thanks to corollary \ref{cor:parallel_removal_v2}, we can remove faces with same direction and orientation in parallel (l.~\ref{pc:08}), while guaranteeing topology preservation.

\begin{algorithm}[h!]
\label{algo::pardircollapse_v0}
\SetKwFor{ForAll}{for all}{do}{end}
\caption[ParDirCollapse]{$ParDirCollapse(X,W,\ell)$}
\KwData{A cubical complex $\complex{n}{X}$, a subcomplex
  $\subcomplex{W}{X}$ which represents faces of $X$ which should not
  be removed, and $\ell \in \Nset$, the number of layers of free faces
  which should be removed from $X$} 
\KwResult{A cubical complex}
\lnl{pc:01}\While{there exists free faces in $X \setminus W$ and $\ell > 0$}{
\lnl{pc:02}  $L=\border{X}^-$\;
\lnl{pc:03}  \For{$t=0 \rightarrow n-1$}{
\lnl{pc:04}    \For{$s=0 \rightarrow 1$}{
\lnl{pc:05}      \For{$d=n \rightarrow 1$}{
\lnl{pc:06}        $E=\{(f,g)$ free for $X \suchthat g \notin W,$\\
                      $\;$~~~~~~~$\type{f,g}=t,$ $\orient{f,g}=s,$ $\mydim{f}=d\}$\;
\lnl{pc:07}        $G=\{(f,g) \in E \suchthat f \in L$ and $g \in L\}$\;
\lnl{pc:08}        $X=X \setminus G$\;
                 }
               }
             }
\lnl{pc:09}  $l = l-1$\;
	   }
\lnl{pc:10} \Return $X$\;
\end{algorithm}

@@

For example, if the input is a corridor (the void of the corridor), then the skeleton should be a line following the main direction of the corridor. 
However, Alg.~\ref{algo::pardircollapse_v0} does not necessarily preserve geometrical features of the input object in the resulting skeleton, which end up being a single vertex.
Moreover, when one wants to preserve geometrical information from the original shape in the skeleton, then it is important to not retain "too much information" during thinning in order to avoid noisy branches in the skeleton.

@@

The lifespan of a face $f$ of $X$ indicates how many "rounds" this face "survives" as a facet in $X$, when removing free pairs with Alg.~\ref{algo::pardircollapse_v0}, and is a good indicator of how important a face can be in an object. Typically, the higher the lifespan is, and the more representative of an object's geometrical feature the face is. The lifespan, sometimes called \myem{saliency}, was used in \cite{Liu2009} (with the name "medial persistence") in order to propose a thinning algorithm in cubical complexes based on two parameters.

@@

Given $S \subset \Zset^n$, the value of $\DD1{S}(x)$ of every $x \in S$ can be naively computed by performing successive morphological dilations of values of the map $\Dist1{S}$. A linear algorithm for computing the map $\DD1{S}$ (with regards to the size of the input image) was proposed in \ref{Chaussard_phd}, and will be explored further in details in a future paper.

@@

@@

\bigskip
\begin{math}
\cplDist1{Y}(f) = \left\{
\begin{array}{ll}
\Dist1{S}(\V2C^{-1}(f)) & \text{if $f$ is an $n$-face}\\
\displaystyle \max_{f \in \cellstrict{g} \inter \subfaces{n}{n}} \cplDist1{Y}(g) & \text{otherwise}
\end{array}
\right.
\end{math}

@@

@@

>>>>

\def\figpath{./images}

\def\figpath{images}
\def\myincludegraphics#1#2{\includegraphics[width=#1]{\figpath/#2}}

@@

\begin{abstract}
We present a novel 3d curvilinear skeletonization algorithm which produces filtered skeletons without needing any user input, thanks to a new parallel algorithm based on the cubical complex framework. We then propose a new path tracing algorithm which uses such curvilinear skeleton in order to produce less noisy images in less time than the classical approach.
\end{abstract}

@@

On a single execution of the main loop of Alg.~\ref{algo::pardircollapse_v0}, only faces located on the border of the complex are removed (l.~\ref{pc:07}). Thanks to corollary \ref{cor:parallel_removal_v2}, we can remove faces with same direction and orientation in parallel (l.~\ref{pc:08}), while guaranteeing topology preservation. Figure \ref{fig::example_algo} depicts the first steps of the algorithm.

\begin{algorithm}[ht!]
\label{algo::pardircollapse_v0}
\SetKwFor{ForAll}{for all}{do}{end}
\caption[ParDirCollapse]{$ParDirCollapse(X,W,\ell)$}
\KwData{A cubical complex $\complex{n}{X}$, a subcomplex
  $\subcomplex{W}{X}$ which represents faces of $X$ which should not
  be removed, and $\ell \in \Nset$, the number of layers of free faces
  which should be removed from $X$} 
\KwResult{A cubical complex}
\lnl{pc:01}\While{there exists free faces in $X \setminus W$ and $\ell > 0$}{
\lnl{pc:02}  $L=\border{X}^-$\;
\lnl{pc:03}  \For{$t=0 \rightarrow n-1$}{
\lnl{pc:04}    \For{$s=0 \rightarrow 1$}{
\lnl{pc:05}      \For{$d=n \rightarrow 1$}{
\lnl{pc:06}        $E=\{(f,g)$ free for $X \suchthat g \notin W,$\\
                      $\;$~~~~~~~$\type{f,g}=t,$ $\orient{f,g}=s,$ $\mydim{f}=d\}$\;
\lnl{pc:07}        $G=\{(f,g) \in E \suchthat f \in L$ and $g \in L\}$\;
\lnl{pc:08}        $X=X \setminus G$\;
                 }
               }
             }
\lnl{pc:09}  $l = l-1$\;
	   }
\lnl{pc:10} \Return $X$\;
\end{algorithm}

\begin{figure}[htp]
\begin{center}
\begin{tabular}{ccccc}
	\myincludegraphics{0.19\textwidth}{invader.pdf} &
	\myincludegraphics{0.19\textwidth}{invader_step1_2.pdf} &
	\myincludegraphics{0.19\textwidth}{invader_step2_2.pdf} &
	\myincludegraphics{0.19\textwidth}{invader_step3.pdf} &
	\myincludegraphics{0.19\textwidth}{invader_step4.pdf} \\
\end{tabular}
\caption[Four first iterations of Alg.~\ref{algo::pardircollapse_v0}]{\label{fig::example_algo} \textbf{Four first iterations of Alg.~\ref{algo::pardircollapse_v0}} running on the left-most shape.}
\end{center}
\end{figure}

@@

For example, if the input is a corridor (the void of the corridor), then the skeleton should be a line following the main direction of the corridor. 
However, Alg.~\ref{algo::pardircollapse_v0} does not necessarily preserve geometrical features of the input object in the resulting skeleton, which end up being a single vertex.
Moreover, when one wants to preserve geometrical information from the original shape in the skeleton, then it is important to not retain \quotes{too much information} during thinning in order to avoid noisy branches in the skeleton.

@@

The lifespan of a face $f$ of $X$ indicates how many \quotes{rounds} this face \quotes{survives} as a facet in $X$, when removing free pairs with Alg.~\ref{algo::pardircollapse_v0}, and is a good indicator of how important a face can be in an object. Typically, the higher the lifespan is, and the more representative of an object's geometrical feature the face is. The lifespan, sometimes called \myem{saliency}, was used in \cite{Liu2009} (with the name \quotes{medial persistence}) in order to propose a thinning algorithm in cubical complexes based on two parameters.

@@

Given $S \subset \Zset^n$, the value of $\DD1{S}(x)$ of every $x \in S$ can be naively computed by performing successive morphological dilations of values of the map $\Dist1{S}$. A linear algorithm for computing the map $\DD1{S}$ (with regards to the size of the input image) was proposed in \cite{Chaussard_phd}, and will be explored further in details in a future paper.

@@

An exemple of these maps is shown on Fig.~\ref{fig::globalopening}.

@@

\bigskip
\begin{math}
\cplDist1{Y}(f) = \left\{
\begin{array}{ll}
\Dist1{S}(\V2C^{-1}(f)) & \text{if $f$ is an $n$-face}\\
\displaystyle \max_{f \in (\cellstrict{g} \inter \subfaces{n}{n})} \cplDist1{Y}(g) & \text{otherwise}
\end{array}
\right.
\end{math}
\bigskip

@@

\begin{figure}[htp]
\begin{center}
\begin{tabular}{cccc}
	\myincludegraphics{0.22\textwidth}{fish.pdf} &
	\myincludegraphics{0.22\textwidth}{fishd.pdf} &
	\myincludegraphics{0.22\textwidth}{fishdd.pdf} &
	\myincludegraphics{0.22\textwidth}{fishcenter.pdf} \\
\end{tabular}
\caption[Example of opening and decenterness map]{\label{fig::globalopening} \textbf{Example of opening and decenterness map - } From left to right: a shape $S \subset \Zset^2$ (in gray), $\Dist1{S}$, $\DD1{S}$ and $\decent{S}$ (low values have dark colour).}
\end{center}
\end{figure}

@@

\subsection{Results}

Algorithm \ref{algo::3d_curvskel} allows to obtain a filtered curvilinear skeleton from a three dimensional complex. The results presented in Fig.~\ref{fig::result_skel} show that the skeletons contain the main geometrical information from the input shapes, and do not contain spurious branches.

\begin{figure}[htp]
\begin{center}
\begin{tabular}{cc|cc}
	\myincludegraphics{0.18\textwidth}{hand2.jpg} &
	\myincludegraphics{0.18\textwidth}{hand_curvskel.jpg} &
	\myincludegraphics{0.24\textwidth}{neptune.jpg} &
	\myincludegraphics{0.24\textwidth}{neptune_curvskel.jpg} \\
\end{tabular}
\caption[Results of algorithm \ref{algo::3d_curvskel}]{\label{fig::result_skel} \textbf{Results of algorithm \ref{algo::3d_curvskel}} for two shapes: a hand (left) and a statue (right). In each pair, the rightmost image represents the skeleton.}
\end{center}
\end{figure}
